---
title: "Lesson3"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Exploring Categorical variables**

### **Frequency table and bar plot**

```{r n, fig.align='center'}
table(mtcars$cyl)
barplot(table(mtcars$cyl))
```

*barplot* different form *histogram*

### **Contingency table**

```{r dd, fig.align='center'}
data(UCBAdmissions)
DF <- as.data.frame(UCBAdmissions)
xt <- xtabs(Freq ~ Gender + Admit,data=DF) # CROSS TAB
xt
```

### **segmented barplot**

```{r dd4, fig.align='center'}
barplot(xt)
```

Visualizing conditional frequency distribution 

the alternative is calculate relative frequency segment barplot

### **Mosaicplot**

### **side-by-side boxplot**

```{r da, fig.align='center'}
data("airquality")
library(dplyr)
airquality <- mutate(airquality, Month = factor(Month)) 
boxplot(Ozone ~ Month, airquality, xlab = "Month", ylab = "Ozone (ppb)")
```

## **Introduction to inference**

* Start with the null hypothesis (H0) that represents the status quo  
* Set an alternative hypothesis (H1) that represents the research question   
* Conduct a hypothesis test under the assumption that the null hypothesis is true, either via simulation or theoretical methods
* If the test results suggest the data do not provide convincing evidence for the alternative hypothesis stick with the null hypothesis
* If they do, reject the null hypothesis in favor of the alternative

**Simulation**

in order to get to the conclusion, we can compare the actual data to a simulated distribution that suppose independence of the variables.

If the simulations do not look like the data ---> reject the null hypothesis



